---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# enfold

<!-- badges: start -->
<!-- badges: end -->

`enfold` is a model ensembling toolbox that makes nested cross-validation
convenient and accessible.  Applicable to all supervised learning
problems, but designed with nuisance model estimation for causal
inference in mind. Supports many types of advanced learner objects found in
modern machine learning ecosystems, such as pipelines and
hyperparameter optimizers. There are three distinctive features of `enfold`:

- List learners, which train one learner and output a list of valid predictions
(e.g., fitting an entire elastic net regularization path)
- Predictions from cross-validated ensembles as first-class feature
- Very simple learner creation via built-in factory makers

## Installation

You can install the development version of `enfold` from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("malik-ck/enfold")
```

## Example

We use the built-in `enfold_demo` data set to predict HbA1c. Suppose we wanted to compare
HbA1c among those in our study that had coronary heart disease to those who did not.
We use a simple mean learner, a random forest, and a generalized linear model (GLM).
For illustration, we create a GLM template in the code block below, though `enfold`
of course ships one as `lrn_glmnet()`.

```{r fit ensemble}
set.seed(7441)
library(enfold)
data(enfold_demo, package = "enfold")

predictors <- enfold_demo[, setdiff(names(enfold_demo), "hba1c")]
outcome <- enfold_demo$hba1c

# Create a GLM template
# fit always needs to be a function of x and y, even if we use more arguments.
# preds always needs to be a function of object and data.
# Additional required arguments are passed as named arguments to make_learner_factory.
# Here, we add "family" without value, which makes the family argument of our learner
# have no default.
my_glm_maker <- make_learner_factory(
  fit = function(x, y) glm(y ~ ., data = data.frame(y = y, x), family = family),
  preds = function(object, data) {
    predict(object, newdata = data, type = "response")
  },
  family
)

# Now instantiate a learner
# The first argument is 'name', added by the learner factory maker, and required
my_glm <- my_glm_maker(name = "My GLM", family = gaussian())

# Other learners to add: mean learner, random forest
mean_learner <- lrn_mean("Mean")
ranger_learner <- lrn_ranger("Random Forest", num.trees = 300)

# Build superlearner ensembles via three-fold cross-validation.
# Do this in an outer three-fold cross-validation loop
# to get pure out-of-fold predictions.
superlearner_ensembles <- initialize_enfold(predictors, outcome) |>
  add_learners(my_glm, mean_learner, ranger_learner) |>
  add_metalearners(mtl_superlearner("SL")) |>
  add_cv_folds(inner_cv = 3L, outer_cv = 3L) |>
  fit()
```

We have a number of functions available to evaluate fit ensembles.
Below, we use `risk()` to assess mean loss for our superlearner
as well as individual learners via `risk_learners()`. Passing
`type = "cv"` calculates risk exclusively on out-of-fold data
using the fold specification added by `add_cv_folds`.

```{r eval ensemble}
risk(superlearner_ensembles, type = "cv", loss_fun = loss_gaussian())

risk_learners(
  superlearner_ensembles,
  type = "cv",
  loss_fun = loss_gaussian()
)
```

We now want to predict HbA1c, setting coronary heart disease to
`1` or `0` throughout our sample. This does not estimate a causal quantity
in this case, but helps us assess whether we expect people with coronary
heart disease to have a different HbA1c, averaged over the covariate
distribution in our study. Note that the data in `enfold_demo` is
simulated and does not provide clinical insight. The predictions we
generate here are cross-validated. We take the mean of predictions to
summarize our results.

```{r cv predict}
# Setting to no CHD
data_no_chd <- predictors
data_no_chd$coronary_heart_disease <- 0

# Setting to all CHD
data_all_chd <- predictors
data_all_chd$coronary_heart_disease <- 1

# Predicting
no_chd_preds <- predict(
  superlearner_ensembles,
  newdata = data_no_chd,
  type = "cv"
)
all_chd_preds <- predict(
  superlearner_ensembles,
  newdata = data_all_chd,
  type = "cv"
)

# Looking at the results...

# Mean among no CHD:
mean(no_chd_preds)

# Mean among all CHD:
mean(all_chd_preds)

# Mean difference 1 - 0:
mean(all_chd_preds) - mean(no_chd_preds)
```

If we wanted non-cross-validated predictions (typical for,
say, model deployment), we would set `outer_cv = NA`
and change `type` to `ensemble`.

Check the package vignettes for advanced functionality,
such as pipelines and grids.

## List learner illustration

To show list learner functionality, we use `enfold`
below to fit one elastic net learner without a metalearner.
When calling `risk_learners`, we can see that `enfold`
treats the elastic net as one learner per lambda.

```{r enfold list}
# Slight quirk: Need to pre-create lambda sequence
make_lambdas <- make_lambda_sequence(
  x = predictors,
  y = outcome,
  nlambda = 10L
)

# Now fit...
elastic_net_illustration <- initialize_enfold(x = predictors, y = outcome) |>
  add_learners(lrn_glmnet(
    "Elnet",
    family = "gaussian",
    lambda = make_lambdas
  )) |>
  add_cv_folds(inner_cv = NA, outer_cv = 5L) |>
  fit()

# ...and estimate risk across the lambda sequence
risk_learners(elastic_net_illustration, type = "cv", loss_fun = loss_gaussian())
```