---
title: "Getting started with enfold"
vignette: >
  %\VignetteIndexEntry{Getting started with enfold}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{quarto::html}
format:
  html:
    toc: true
    toc-depth: 3
knitr:
  opts_chunk:
    collapse: true
    comment: "#>"
editor: 
  markdown: 
    wrap: 72
---

```{r setup, include = FALSE}
library(enfold)
```

## Overview

`enfold` (*ensemble fold*) is a package that makes model stacking and
ensembling in R quick and convenient. Users have very fine-grained
control over *how* ensembles are built, and ensembles can be
cross-validated with ease.

`enfold` is designed to be as type-agnostic as is reasonable. This means
that the package internally does nothing to change the data you hand in.
It is entirely the job of the learners to take training data, to know
how to fit it, and to predict on validation data. This is very helpful
for any user that is interested in implementing their own learner: this
is easy in `enfold`, and users do not need to be scared of any hidden
effects the package has on their data. For users that do not need this
flexibility, `enfold` implements many learner templates that are
well-equipped to handle all types of matrices and data frames as
predictor input.

`enfold` was developed as a companion package to a package for causal
inference via target trial emulation named `enact`, and its set of
features somewhat reflects this. For example, getting cross-validated
predictions on mutated data sets is a first-class feature and
recommended whenever ensembling is done for a doubly robust estimator.
Users that use `enfold` for pure prediction tasks will still find many
advanced features, such as pipelines and hyperparameter optimization,
but might find it lacking for model deployment.

This guide will use the `Boston` housing data set to introduce the core
features and workflow of the `enfold` package. At the end, users should
know how to estimate ensembles in `enfold` using provided templates, as
well as some of the quirks the package has. Following guides will go
into detail on some advanced functionality, such as pipelines, grids,
constructors for custom learners and hyperparameter optimizers, and the
package's internal structure.

## Data

We use a simulated data set from the `enfold` package named
`enfold_demo`. The outcome we are interested in predicting is `hba1c`,
which reflects a person's average blood glucose level over roughly the
preceding two to three months. It's used to both diagnose diabetes and
to monitor glycemic control in people already diagnosed.

```{r data}
data("enfold_demo", package = "enfold")

x <- enfold_demo[, setdiff(names(Boston), "hba1c")]
y <- enfold_demo$hba1c

cat("Rows:", nrow(x), "| Predictors:", ncol(x), "\n")
```

Using enfold now consists of five steps. First, we initialize a task by
passing data. Following this, we tell the task which learner(s) we want
to use, which metalearner(s) (ensembling algorithms) we want to use, and
how we want to cross-validate, in any order. Last, we call fit() on the
task to fit the ensemble(s). Afterwards, a suite of evaluator functions
will tell us how well our ensembles and learners did. Of course, we can
also call `predict` on a fitted task to get predictions for all
ensembles.

In this guide, we will go in the order: initialize, cross-validation,
learners, metalearners, fit, and then investigate the performance of our
ensembles. Note that, when working on your own projects, you will
probably want to explore your data before analyses, say by plotting
variables, gaining some understanding of how they are distributed, and
similar. We omit this for brevity.

## Step 1 — Initialize a task

`initialize_enfold()` creates an `enfold_task` object that stores the
predictor data `x` and outcome `y`.

::: {.callout-tip collapse="true" title="Advanced note: data storage"}
After initialization, the task does not actually store `x` and `y`. It
instead puts a reference to the handed-in data into a fresh environment
and locks it. This ensures that data is not unnecessarily copied. If
users delete `x` and `y` from their global environment, they are still
referenced from this environment and therefore available.

`enfold` has a very piping-friendly workflow, which might not play well
with the reference semantics that environments introduce. This issue
should be alleviated by ensuring that the data stored in `x` and `y` are
never changed by the `enfold` task (only by learners that receive a copy
of the data for fitting). Since their environment is locked, this can
also not happen by accident.
:::

We can afterwards look at the slots inside the task:

```{r initialize}
task <- initialize_enfold(x, y)
task
```

That's many empty slots! We will fill these as we continue adding bits
to the task. By the time the task is fully specified but not yet fit,
the only remaining empty slots will be `fit_objects` and `ensembles`,
which hold fitted learners and metalearners.

## Step 2 — Add learners

Learners are added via `add_learners()`. The learner templates provided
by `enfold` follow the naming convention that they start with `lrn_*`.
Each learner template has a `name` argument which needs to be filled, as
well as (optionally) additional arguments for *hyperparameters*. This is
how we might fit random forests with different values for, say, *mtry*.
You can check each learner's documentation for more detail.

`add_learners()` takes as arguments `task` and dots, which means that
its first argument needs to be a task object, and each following
argument is a learner object. The more natural way of calling
`add_learners()` is via *piping*. We will use pipes throughout this
guide; users not familiar with piping can expand the optional code block
below, where both options are shown.

To start, we will add two learners. The first will be the empirical mean
of the outcome (`lrn_mean()`), which is a useful benchmark: it uses
*very* little information to arrive at a prediction. The second will be
a generalized linear model (`lrn_glm()`), and we will tell it to use a
Gaussian (normal) likelihood via the `family` argument.

```{r learners}
task <- task |>
  add_learners(
    lrn_mean("mean"), # Intercept-only benchmark
    lrn_glm("glm", family = gaussian()) # Gaussian GLM
  )
```

::: {.callout-tip collapse="true" title="How to pipe"}
In R and other programming languages, a `pipe` can be used to resolve
nested functions by composing them from left to right. For example, if
we want to take the mean of some variable and then round it, we could
write either `round(mean(my_variable))`, or
`mean(my_variable) |> round()`. While piping is overkill here, it
illustrates the concept: one takes the object created on the left of the
pipe, and inserts it as the first argument to the function on the right.
In this example, we first calculate the mean of `my_variable`, and then
put the result as the first argument into `round`.

In cases where we want to chain many functions together, piping helps
make things more readable. Below you can see how we could add learners
to a task without and with piping; the latter is the expected and
recommended workflow in `enfold`.

```{r pipes}

# Without piping
add_learners_no_pipe <- add_learners(
  task,
  lrn_mean("mean"),
  lrn_glm("glm", family = gaussian())
)

# With piping
add_learners_with_pipe <- task |>
  add_learners(
    lrn_mean("mean"),
    lrn_glm("glm", family = gaussian())
  )

#...or to do it all in one shot without having an initialization object sitting around idly:
initialize_and_add_learners <- initialize_enfold(x, y) |>
  add_learners(
    lrn_mean("mean"),
    lrn_glm("glm", family = gaussian())
  )

```
:::

You can use a host of other learner templates if you prefer. Available built-in 
learner constructors include `lrn_mean`, `lrn_glm`, `lrn_gam`, `lrn_glmnet`, 
`lrn_ranger`, `lrn_xgboost`, and more. See their help pages for details.

## Step 3 — Add metalearners

Metalearners take the out-of-fold predictions of all base learners as
input and learn a combination rule. They are added with
`add_metalearners()` and follow the `mtl_*` naming convention.

```{r metalearners}
task <- task |>
  add_metalearners(
    mtl_selector("selector"), # Picks the single best base learner
    mtl_superlearner("superlearner") # Learns an optimal convex combination
  )
```

`mtl_selector` is a discrete rule that assigns all weight to the base
learner with the lowest cross-validated loss. `mtl_superlearner` instead builds a
convex combination of learners. To be specific, it regresses the outcome
on cross-validated predictions of each learner under the constraints that no
learner can have a negative coefficient, and that the sum of coefficients is `1`.
It is a good default metalearner to use and recommended over `mtl_selector`
for performance. Both accept a `loss_fun` argument to control which
loss drives the weighting (default: `loss_gaussian()`).

This is what we refer to as ensemble building: we take cross-validated predictions
of a number of base learners, and try to find some way of combining them to
optimize performance. The most common metalearner is simple choosing (which
`mtl_selector()` does), but there is nothing wrong with using more complex
metalearners (like `mtl_superlearner()`). However, with very complex metalearners
and many base learners we risk overfitting. While the superlearner is constrained
enough to rarely ever overfit, having `mtl_selector()` as a simpler benchmark
and cross-validating ensemble building can make us a little more sure.


## Step 4 — Add cross-validation folds

`add_cv_folds()` specifies how folds are created.

-   `inner_cv` controls the folds used to build metalearners (the inner
    loop).
-   `outer_cv` controls the folds used to evaluate the full ensemble
    (the outer loop).

Pass an integer for simple V-fold CV, a custom function
`function(n, ...)`, or `NA` / `NULL` to skip that level.

```{r folds}
task <- task |>
  add_cv_folds(inner_cv = 5L, outer_cv = 5L)
```

Alternatively you can build an `enfold_cv` object separately with
`create_cv_folds()` and pass it via the `cv` argument. This is useful
when you want to reuse the same fold splits across multiple tasks.

```{r folds-reuse, eval = FALSE}
cv <- create_cv_folds(n = nrow(x), inner_cv = 5L, outer_cv = 5L)
task_alt <- initialize_enfold(x, y) |>
  add_learners(lrn_glm("glm", family = gaussian())) |>
  add_metalearners(mtl_selector("selector")) |>
  add_cv_folds(cv = cv)
```

## Step 5 — Fit

`fit()` trains all learners and metalearners according to the fold
structure.

```{r fit}
task <- fit(task)
task
```

For parallel execution across outer folds, set a `future` plan *before*
calling `fit()`:

```{r parallel, eval = FALSE}
future::plan(future::multisession, workers = 4L)
task <- fit(task)
future::plan(future::sequential)   # reset afterwards
```

## Step 6 — Evaluate

### Cross-validated predictions

`predict(task, type = "cv")` returns out-of-fold ensemble predictions
stitched back together in the original observation order. When multiple
metalearners are present, a named list is returned.

```{r predict-cv}
cv_preds <- predict(task, type = "cv")
```

### Overall risk

`risk()` summarizes cross-validated performance as mean loss per
metalearner.

```{r risk}
risk(task, loss_fun = loss_gaussian(), type = "cv")
```

### Per-fold risk

`fold_risk()` breaks the same summary down by outer fold, which is
useful for assessing stability.

```{r fold-risk}
fold_risk(task, loss_fun = loss_gaussian())
```

### Observation-level loss

`cv_loss()` returns a data frame with one loss value per observation,
enabling loss distribution plots and outlier inspection.

```{r cv-loss}
obs_loss <- cv_loss(task, loss_fun = loss_gaussian())
head(obs_loss)
```

### Base-learner predictions and risk

`predict_learners()` and `risk_learners()` expose the individual
base-learner predictions before any metalearning, which helps to
understand which learners contribute most.

```{r learner-risk}
risk_learners(task, loss_fun = loss_gaussian(), type = "cv")
```

### Predicting on new data

`predict(task, type = "ensemble")` uses a fitted ensemble (not
cross-validated) to score new observations. For a cross-validated task
you must specify which outer-fold ensemble to use via
`ensemble_fold_id`.

```{r predict-new, eval = FALSE}
new_preds <- predict(task, newdata = x[1:10, ],
                     type = "ensemble", ensemble_fold_id = 1L)
```

## Advanced functionality

You now know how to fit and evaluate ensembles via the `enfold` package!
If you want, you can take a deep dive into some of the more advanced
features the package has. Below we will briefly introduce pipelines and
grids, which are commonly available in machine learning libraries and
allow for very flexible modeling. Both of them will also receive
separate guides that go into more detail for interested users.

### Pipelines

`make_pipeline()` chains learners so that each stage's output feeds into
the next. This is useful for screeners, basis expanders, or
post-processing steps. Branching within a stage creates one independent
path per combination.

```{r pipelines, eval = FALSE}
# A screener upstream, then two competing learners downstream
pipe <- make_pipeline(
  scr_correlation("screener", cutoff = 0.1),
  list(
    lrn_glm("glm",  family = gaussian()),
    lrn_mean("mean")
  )
)

task_pipe <- initialize_enfold(x, y) |>
  add_learners(pipe) |>
  add_metalearners(mtl_selector("selector")) |>
  add_cv_folds(inner_cv = 5L, outer_cv = 5L) |>
  fit()
```

`predict()` on a fitted pipeline returns a named list with one entry per
path (e.g. `"screener/glm"`, `"screener/mean"`). Metalearners receive
each path as an independent learner.

### Hyperparameter grids

`make_grid()` cross-validates a learner over a grid of hyperparameter
combinations. It requires `inner_cv` to be non-`NULL`.

```{r grids, eval = FALSE}
grid <- make_grid(
  name_prefix      = "glmnet",
  lrn_glmnet,
  alpha            = c(0, 0.5, 1),
  lambda           = c(0.01, 0.1, 1)
)

task_grid <- initialize_enfold(x, y) |>
  add_learners(grid) |>
  add_metalearners(mtl_superlearner("sl")) |>
  add_cv_folds(inner_cv = 5L, outer_cv = 5L) |>
  fit()
```

Each hyperparameter combination is treated as a separate learner and
handed to the metalearner together with the other base learners.

## Session info

```{r session-info}
sessionInfo()
```
