---
title: "Getting started with enfold"
vignette: >
  %\VignetteIndexEntry{Getting started with enfold}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{quarto::html}
format:
 html: 
    toc: true
    toc-depth: 3
knitr:
  opts_chunk:
    collapse: true
    comment: "#>"
editor: 
  markdown: 
    wrap: 72
---

```{r setup, include = FALSE}
set.seed(4112)
library(enfold)
```

## Overview

`enfold` (*ensemble fold*) is a package that makes model stacking and
ensembling in R quick and convenient. Users have very fine-grained
control over *how* ensembles are built, and ensembles can be
cross-validated with ease. There are two use cases with which in mind
`enfold` was developed:

- The use and evaluation of bespoke learners which are not implemented
in any machine learning package, where friction-less learner template
creation is important
- Ensemble validation via nested cross-validation, say for nuisance
model estimation in causal inference

`enfold` is designed to be as type-agnostic as is reasonable. This means
that the package internally does nothing to change the data you hand in.
It is entirely the job of the learners to take training data, to know
how to fit it, and to predict on validation data. This is very helpful
for any user that is interested in implementing their own learner: this
is easy in `enfold`, and users do not need to be scared of any hidden
effects the package has on their data. For users that do not need this
flexibility, `enfold` implements many learner templates that are
well-equipped to handle all types of matrices and data frames as
predictor input.

`enfold` was developed as a companion package to a package for causal
inference via target trial emulation named `enact`, and its set of
features somewhat reflects this. For example, getting cross-validated
predictions on mutated data sets is a first-class feature and
recommended whenever ensembling is done for a doubly robust estimator.
Users that use `enfold` for pure prediction tasks will still find many
advanced features, such as pipelines and hyperparameter optimization,
but will find it lacking for model deployment.

This guide will use a simulated medical data set to introduce the core
features and workflow of the `enfold` package. At the end, users should
know how to estimate ensembles in `enfold` using provided templates, as
well as some of the quirks the package has. Following guides will go
into detail on some advanced functionality, such as pipelines, grids,
constructors for custom learners and hyperparameter optimizers, and the
package's internal structure.

## Data

We use a simulated data set from the `enfold` package named
`enfold_demo`. The outcome we are interested in predicting is `hba1c`.
This is a common blood test result that reflects a person's average 
blood glucose level over roughly the preceding two to three months. 
It's used to both diagnose diabetes and to monitor glycemic control 
in people already diagnosed. We want to predict `hba1c` from a number
of demographic and medical predictors seen below:

```{r data}
data("enfold_demo", package = "enfold")

predictors <- enfold_demo[, setdiff(names(enfold_demo), "hba1c")]
outcome <- enfold_demo$hba1c

cat("Rows:", nrow(predictors), "| Predictors:", ncol(predictors), "\n")
cat("Predictor names:", paste(colnames(predictors), collapse = ", "), "\n")
```

Using enfold now consists of five steps. First, we initialize a task by
passing data. Following this, we tell the task which learner(s) we want
to use, which metalearner(s) (ensembling algorithms) we want to use, and
how we want to cross-validate, in any order. Last, we call fit() on the
task to fit the ensemble(s). Afterwards, a suite of evaluator functions
will tell us how well our ensembles and learners did. Of course, we can
also call `predict` on a fitted task to get predictions for all
ensembles.

In this guide, we will go in the order: initialize, learners, metalearners, 
cross-validate, fit, and then investigate the performance of our
ensembles. Note that, when working on your own projects, you will
probably want to explore your data before analyses. We omit this for brevity;
the data are simulated such that they are well-behaved.

## Step 1 — Initialize a task

`initialize_enfold()` creates an `enfold_task` object that stores the
predictor data `x` and outcome `y`.

::: {.callout-tip collapse="true" title="Advanced note: data storage"}
After initialization, the task does not actually store `x` and `y`. It
instead puts a reference to the handed-in data into a fresh environment
and locks it. This ensures that data is not unnecessarily copied. If
users delete `x` and `y` from their global environment, they are still
referenced from this environment and therefore available.

`enfold` has a very piping-friendly workflow, which might not play well
with the reference semantics that environments introduce. This issue
should be alleviated by ensuring that the data stored in `x` and `y` are
never changed by the `enfold` task (only by learners that receive a copy
of the data for fitting). Since their environment is locked, this can
also not happen by accident.

If you have large data or little memory, enfold has support for file-backed
systems via the `bigstatsr` and `arrow` packages. Check the big data guide
for more details.
:::

We can afterwards look at the slots inside the task:

```{r initialize}
task <- initialize_enfold(x = predictors, y = outcome)
task
```

That's many empty slots! We will fill these as we continue adding bits
to the task. By the time the task is fully specified but not yet fit,
the only remaining empty slots will be `fit_objects` and `ensembles`,
which hold fitted learners and metalearners.

## Step 2 — Add learners

Learners are added via `add_learners()`. The learner templates provided
by `enfold` follow the naming convention that they start with `lrn_*`.
Each learner template has a `name` argument which needs to be filled, as
well as (optionally) additional arguments for *hyperparameters*. This is
how we might fit random forests with different values for, say, *mtry*.
You can check each learner's documentation for more detail.

`add_learners()` takes as arguments `task` and dots, which means that
its first argument needs to be a task object, and each following
argument is a learner object. The more natural way of calling
`add_learners()` is via *piping*. We will use pipes throughout this
guide; users not familiar with piping can expand the optional code block
below, where both options are shown.

To start, we will add two learners. The first will be the empirical mean
of the outcome (`lrn_mean()`), which is a useful benchmark: it uses
*very* little information to arrive at a prediction, and any other learner
should beat it comfortably. The second will be a generalized linear model 
(`lrn_glm()`), and we will tell it to use a Gaussian (normal) likelihood via 
the `family` argument.

```{r learners}
task <- task |>
  add_learners(
    lrn_mean("mean"), # Intercept-only benchmark
    lrn_glm("glm", family = gaussian()) # Gaussian GLM
  )
```

::: {.callout-tip collapse="true" title="How to pipe"}
In R and other programming languages, a `pipe` can be used to resolve
nested functions by composing them from left to right. For example, if
we want to take the mean of some variable and then round it, we could
write either `round(mean(my_variable))`, or
`mean(my_variable) |> round()`. While piping is overkill for this example, it
illustrates the concept: one takes the object created on the left of the
pipe, and inserts it as the first argument to the function on the right.
In this example, we first calculate the mean of `my_variable`, and then
put the result as the first argument into `round`.

In cases where we want to chain many functions together, piping helps
make things more readable. Below you can see how we could add learners
to a task without and with piping; the latter is the expected and
recommended workflow in `enfold`.

```{r pipes}
# Without piping
add_learners_no_pipe <- add_learners(
  task,
  lrn_mean("Mean"),
  lrn_glm("GLM", family = gaussian())
)

# With piping
add_learners_with_pipe <- task |>
  add_learners(
    lrn_mean("Mean"),
    lrn_glm("GLM", family = gaussian())
  )

#...or to do it all in one shot without having an initialization object sitting around idly:
initialize_and_add_learners <- initialize_enfold(x = predictors, y = outcome) |>
  add_learners(
    lrn_mean("Mean"),
    lrn_glm("GLM", family = gaussian())
  )
```
:::

You can use a host of other learner templates if you prefer. Available built-in 
learner constructors include `lrn_mean`, `lrn_glm`, `lrn_gam`, `lrn_glmnet`, 
`lrn_ranger`, `lrn_xgboost`, and more. See their help pages for details.

## Step 3 — Add metalearners

Metalearners take the out-of-fold predictions of all base learners as
input and learn a combination rule. They are added with
`add_metalearners()` and their templates follow the `mtl_*` naming convention.

```{r metalearners}
task <- task |>
  add_metalearners(
    mtl_selector("Selector"), # Picks the single best base learner
    mtl_superlearner("Superlearner") # Learns an optimal convex combination
  )
```

`mtl_selector` is a discrete rule that assigns all weight to the base
learner with the lowest cross-validated loss. `mtl_superlearner` instead builds a
convex combination of learners. To be specific, it regresses the outcome
on cross-validated predictions of each learner and optimizes under the constraints that no
learner has a negative weight, and that the sum of weights is `1`.
It is a good default metalearner to use and recommended over `mtl_selector`
for performance. Both accept a `loss_fun` argument to control which
loss drives the weighting (default: `loss_gaussian()`).

This is what we refer to as ensemble building: we take cross-validated predictions
of a number of base learners, and try to find some way of combining them to
optimize performance. The most common metalearner is simple choosing (which
`mtl_selector()` does), but there is nothing wrong with using more complex
metalearners (like `mtl_superlearner()`). However, if our metalearners are very
complex and we have many learners, we risk overfitting. The superlearner is
still simple enough for the overfitting risk to be very small, but to be sure
we can cross-validate our ensemble building process and compare its performance
to the even simpler `mtl_selector()`.


## Step 4 — Add cross-validation folds

To add cross-validation folds, we use the function `add_cv_folds()`.
Besides the task object, it takes one value for `inner_cv` and one for `outer_cv`.
Each controls a different part of the cross-validation process. Note that
it is allowed to make one of them `NA`, but not both. The simplest way to
instruct `add_cv_folds()` is to enter an integer or `NA` to each argument, 
which enables simple V-fold cross-validation. Here is how they compare:

- `outer_cv` controls validation for the full ensemble. For example,
passing `outer_cv = 5` means that we build our ensembles five times in evenly-sized
training sets, and each observation acts as validation for one of the training sets. 
This is useful to get an honest performance evaluation of our
metalearners, or to generate out-of-fold predictions on mutated data sets.
If we set this to `NA`, only one ensemble is built on the full data.
- `inner_cv` controls how ensembles are built. For example, passing `inner_cv = 5`
means that five cross-validation folds are used to build each ensemble. If this
is `NA`, no ensemble is built. Instead, learners are trained without any metalearning 
and provide cross-validated predictions via as many folds as specified in `outer_cv`.

::: {.callout-tip collapse="true" title="Advanced note: how inner and outer cross-validation compose"}
The procedure underlying this is also referred to as *nested cross-validation*.
In words, if both `inner_cv` and `outer_cv` have a non-`NA`
value, the folds for `inner_cv` are nested within each training set in `outer_cv`.
This means that an ensemble is built entirely within one of the training sets
in `outer_cv`: 

1. the training set is treated as a full data-set, 
2. folds are created for that training set, 
3. the ensemble is built using cross-validated predictions for each learner,
4. learners are re-trained on the full training set,
5. ensemble weights and learner predictions are used to generate out-of-fold
predictions on the outer validation set.

This is repeated for each outer fold, such that each metalearner has fully
cross-validated predictions that can be used as-is or to compare metalearner performance.

:::

For this vignette, we will want to compare the performance of our metalearners.
Therefore, we do both inner and outer cross-validation:

```{r folds}
task <- task |>
  add_cv_folds(inner_cv = 5L, outer_cv = 5L)
```

Alternatively we could build an `enfold_cv` object separately with
`create_cv_folds()` and pass it via the `cv` argument. This is useful
if we want to reuse the same fold splits across multiple tasks.

```{r folds-reuse, eval = FALSE}
separate_cv <- create_cv_folds(
  n = nrow(predictors),
  inner_cv = 5L,
  outer_cv = 5L
)

task_separate_cv <- initialize_enfold(x = predictors, y = outcome) |>
  add_cv_folds(cv = separate_cv)
```

To summarize: if we want access to cross-validated predictions of model ensembles, we need to
pass a non-`NA` value to both `inner_cv` and `outer_cv`. If we just want to train
one ensemble and do not care about an honest performance evaluation, we set `outer_cv`
to `NA`. If we do not want to build ensembles and just want cross-validated
predictions for our learners, we set `inner_cv` to `NA`.

::: {.callout-tip collapse="true" title="Advanced note: passing functions for complex CV schemes"}
Passing integers into `inner_cv` and `outer_cv` is fully sufficient if we only
need simple V-fold validation. Unfortunately, we sometimes need more. For
example, if we have clustered data, using clustered cross-validation is recommended:
all observations within a cluster should end up in the same fold.

In such cases, we can instead pass a function to `inner_cv` and `outer_cv`,
which we create via `make_cv_function`. This allows us to use any cross-validation
scheme implemented by us or others. Using functions from the `origami` package is
recommended, because it creates fold objects in the right format: a named list with
two entries, one named `training_set` and one named `validation_set`, 
each containing a vector of indices for the training and validation observations in each fold.

We call `make_cv_function` as follows. The first argument we pass is the cross-validation
function we want to use. Then, we simply pass all other arguments required by the function,
which will often at least be the number of folds `V`. Lastly, if we have any arguments that
are vectors of length `n`, we say which they are by passing their names to the `.subset` argument. 
It is required that the cross-validation function we use has an argument `n`,
and that we don't pass a value to it ourselves.

The code block below shows how we could do cross-validation,
stratified by the variable `coronary_heart_disease`. We remove it as a predictor
and instead pass it as part of our cross-validation instruction:

```{r complex folds}
# Get CHD as separate variable
chd <- enfold_demo$coronary_heart_disease

# Remove it from our predictors
predictors_no_chd <- predictors[,
  -which(colnames(enfold_demo) == "coronary_heart_disease")
]

# Create a new task for illustration that stratifies cross-validation
task_with_stratified_folds <- initialize_enfold(
  x = predictors_no_chd,
  y = outcome
) |>
  add_cv_folds(
    inner_cv = make_cv_function(
      origami::make_folds,
      V = 5L,
      strata_ids = chd,
      .subset = "strata_ids"
    ),
    outer_cv = make_cv_function(
      origami::make_folds,
      V = 5L,
      strata_ids = chd,
      .subset = "strata_ids"
    )
  )
```
:::

## Step 5 — Fit

That's all we need to specify the task! We can now call `fit()`, which fits
our learners and metalearners according to the specifications we have given.

```{r fit}
task <- fit(task)
task
```

`enfold` can also fit in parallel via the packages `future`
and `future.apply`. To do this, we use `future::plan()` to create a parallelization
plan ourselves, and then call `fit()`. 

In `enfold`, parallelization can be done across three levels in total: across outer folds,
across learners, and across inner folds. `enfold` also tries to parallelize in exactly this order.
The code below uses a multisession plan with four workers.

```{r parallel, eval = FALSE}
future::plan(future::multisession, workers = 4L)
task <- fit(task)
future::plan(future::sequential) # reset afterwards
```

Since we had `outer_cv` enabled, `enfold` parallelized over outer folds. If we had set `outer_cv`
to `NA` instead, it would have cross-validated over learners. 

If we have sufficiently many available workers on our machine, we can also set up a nested plan. 
The creation of such a nested plan is shown below. If we called `fit()` after setting up the nested 
plan below, we would parallelize across all three levels: outer folds, learners, and inner folds.

```{r nested parallel, eval = FALSE}
# Nested multisession plan
plan(list(
  tweak(multisession, workers = 2),
  tweak(multisession, workers = 2),
  tweak(multisession, workers = 2)
))

# If task was not yet fit, fit() here

# Reset afterwards
future::plan(future::sequential)
```

When parallelizing, and especially when  doing nested parallelization, be mindful
of your computer's resources. Use `future::availableCores()` to see how many workers you can use.

## Evaluating Ensembles after Fitting

The main functions we can apply to our task after fitting are `predict()`, `risk()`, and `loss()`.
`predict()` uses our ensemble to predict, `risk()` summarizes performance as mean loss over
a user-specified loss function, and `loss()` gives us observation-level loss values. 
Each of these functions has a `type` argument that controls whether we want to use cross-validated 
predictions or predictions from one ensemble. Each also has a version post-fixed by `_learner` that
evaluates all individual learners instead: `predict_learners()`, `risk_learners()`, and `loss_learners()`.

Usually, if we cross-validate ensemble creation, we will use `type = "cv"`.
This will re-use the outer fold specification to predict and calculate loss purely on
out-of-fold data. However, if we want to use our cross-validated ensembles on an entirely
new data set where the outer fold specification from the task does not apply anymore,
we need to choose one of the ensembles to predict with via the `ensemble_fold_id` argument.
A better alternative in such a scenario might be to re-fit the task with `outer_cv = NA`. 
This leaves us with one ensemble trained on the full data, and we can call `type = "ensemble"` 
without ambiguity. Let us go through a few examples.

### Cross-validated model evaluation

If we call `predict()` with `type = "cv"`, we get out-of-fold predictions given the outer folds
from our task. The exact output format is determined by the metalearner used, though the
built-in metalearners pre-fixed with `mtl_*` all output vectors.

```{r predict-cv}
cv_preds <- predict(task, type = "cv")
lapply(cv_preds, head)
```

Attached to the cross-validated predictions is an `indices` attribute. If we used resampling
schemes where one observation might end up in multiple validation folds, this index
tells us to which observation a prediction belongs. In most cases, it will simply count
from `1` to `n`.

We can now calculate cross-validated risk and loss to compare metalearners.
This requires passing a loss function to `loss()` and `risk()`. The default,
`loss_gaussian()`, tends to do well for model comparison.
To get a full picture of ensemble performance, we will also use the corresponding
evaluation functions for learners.

```{r risk and loss}
# Ensemble and learner loss
task_loss <- loss(task, loss_fun = loss_gaussian(), type = "cv")
task_loss_learners <- loss_learners(
  task,
  loss_fun = loss_gaussian(),
  type = "cv"
)

# Ensemble losses:
head(task_loss)

# Learner losses:
head(task_loss_learners)

# Ensemble and learner risks
# Ensemble risks:
risk(task, loss_fun = loss_gaussian(), type = "cv")

# Learner risks:
risk_learners(task, loss_fun = loss_gaussian(), type = "cv")
```

As we can see, the GLM outperformed the mean learner significantly.
The difference in risk between the selector and the metalearner is practically zero. 
This can happen when one learner dominates so much that the superlearner wants to 
assign all weight to it.

There is one last evaluation function, which can only be used for cross-validated ensembles.
`fold_risk()` outputs the risk per fold, which gives an overview over the variability in risk
over folds. 

```{r per fold risk}
fold_risk(task, loss_fun = loss_gaussian())
```

### Single ensemble evaluation

If we set `outer_cv = NA`, all functions besides `fold_risk` still work,
but derive their results from a single ensemble trained on the full data. Using
`loss()` and `risk()` therefore report on loss and risk determined in-sample,
which is not an honest loss evaluation. However, fitting a single ensemble on
full data is typically what we want to do to predict on entirely new data.
Below is a brief example. We estimate an ensemble via a superlearner
and three learners, and predict on new data:

```{r outer off}
one_ensemble_only <- initialize_enfold(x = predictors, y = outcome) |>
  add_learners(
    lrn_mean("Mean"),
    lrn_glm("GLM", family = gaussian()),
    lrn_ranger("Random Forest", num.trees = 200)
  ) |>
  add_metalearners(
    mtl_superlearner("SL", loss_fun = loss_gaussian())
  ) |>
  add_cv_folds(inner_cv = 5L, outer_cv = NA) |>
  fit()

# Superlearner predictions can now be gotten on new data
predict(one_ensemble_only, newdata = predictors[1:5, ], type = "ensemble")
```

## Advanced functionality

You now know how to fit and evaluate ensembles via the `enfold` package!
You can now use `enfold` for your own projects, or read the other guides
to learn about some of the advanced functionality of the package.
As a little teaser, the rest of this guide will briefly introduce list learners,
pipelines, and grids. The former is a feature that `enfold` has and that is
unique to it, while the latter two are important in many modern machine
learning workflows. Taken together, standard learners, list learners,
pipelines, and grids make up all the different types of learners in the `enfold` ecosystem.

### List learners

One key feature of `enfold` is that it can treat the output of one learner as
multiple valid predictions. This is useful in many settings. A common example
is fitting elastic nets via `glmnet`: most packages implement elastic nets
by adding yet another layer of cross-validation to select the optimal regularization
strength, and the trained learner then uses the single best regularization strength
to build ensembles. This is somewhat unfortunate: we build ensembles precisely
to make decisions about which learners can contribute well to our final prediction,
and we might as well treat elastic nets with different regularization strengths as different learners.
It is therefore an option to instead use the entire regularization path to build ensembles,
which saves computational resources and evaluates regularization strength at the
actual ensemble building step.

The code block below shows how this works. We use the `lrn_glmnet` template,
which requires us to pass a regularization sequence ourselves via the `lambda` argument.
We will use a function available through the `enfold` package named `make_lambda_sequence()`,
which takes `x` and `y` as arguments to generate an appropriate regularization sequence.
It also takes `nlambda` as an argument, which controls how many lambda values are generated
(default: 30). It is good to use a generous `nlambda` since this does not increase
computational time. The `glmnet` package defaults to 100, and `enfold` only defaults
to less to prevent nasty memory surprises at predict time. We set `inner_cv = NA`, 
which is how we fit learners that output cross-validated predictions without ensembling.
We use a very small `nlambda` for this guide to make the output more readable.

```{r glmnet show}
# Initialize lambda sequence creation
lambda_seq <- make_lambda_sequence(x = predictors, y = outcome, nlambda = 10L)

glmnet_task <- initialize_enfold(x = predictors, y = outcome) |>
  add_learners(
    lrn_glmnet("LASSO", alpha = 1, lambda = lambda_seq, family = "gaussian")
  ) |>
  add_cv_folds(inner_cv = NA, outer_cv = 5L) |>
  fit()
```

What happens now if we use `predict_learners()` or `risk_learners()`?
Simple: `enfold` knows that `lrn_glmnet` outputs a list of predictions,
so it treats each entry as its separate learner. `risk_learners()` can therefore
tell us what the optimal regularization strength for our data set is.

```{r glmnet risk}
risk_learners(glmnet_task, loss_fun = loss_gaussian(), type = "cv")

# The minimizer only:
setNames(
  min(risk_learners(glmnet_task, loss_fun = loss_gaussian(), type = "cv")),
  names(which.min(risk_learners(
    glmnet_task,
    loss_fun = loss_gaussian(),
    type = "cv"
  )))
)
```

We can alternatively set `outer_cv = NA`, and enable `inner_cv`.
While this prevents us from evaluating cross-validated losses along the regularization path, 
it would build for us ensembles along it to predict on new data.
If we used `mtl_selector()` as metalearner here, we would emulate what `cv.glmnet`
does with `s = "lambda.min"`: calling `predict` on the fitted task outputs only 
the regularization strength deemed optimal by validation data.

There are many templates that are or can be list learners. An important example
besides `lrn_glmnet()` is `lrn_xgboost()`, which can output a list of predictions
after set numbers of boosting rounds.

### Pipelines

`make_pipeline()` chains learners so that each stage's output feeds into
the next. This is useful for screeners, basis expanders, or
post-processing steps. Branching within a stage creates one independent
path per combination. Similar to a list learner, a pipeline learner
outputs a list of valid predictions, in this case one per terminal node.

```{r pipelines, eval = TRUE}
# A screener upstream, then two competing learners downstream
pipe <- make_pipeline(
  scr_svd("Screener", n_components = 4),
  list(
    lrn_glm("GLM", family = gaussian()),
    lrn_mean("Mean")
  )
)

task_pipe <- initialize_enfold(x = predictors, y = outcome) |>
  add_learners(pipe) |>
  add_metalearners(mtl_selector("Selector")) |>
  add_cv_folds(inner_cv = 5L, outer_cv = 5L) |>
  fit()

# Example risk calculation, one per terminal node
risk_learners(task_pipe, loss_fun = loss_gaussian(), type = "cv")
```

### Hyperparameter grids

Grids are how we can create multiple versions of one learner with
different hyperparameters. We can include a grid as a standard learner,
in which case all possible hyperparameter combinations are trained.
We can also wrap a search engine around a grid, which allows us to search
and optimize over admissible hyperparameter combinations. An example
search engine is `grd_early_stop()`, which trains learners from random
hyperparameter draws until validation error stops improving. One unique
part of hyperparameter optimization in `enfold` is that search engines
peek at ensemble building folds. In other words: `enfold` does not set up
another set of cross-validation folds to find an optimal hyperparameter
configuration for a learner. It instead tries to find the combination
that contributes best to the ensemble being built. This treats hyperparameter
optimization as a method to prevent exhaustively searching all learner
combinations. This spares computational resources and performance evaluation
is still honest if `outer_cv` is not `NA`.

To define a grid, we use the function `specify_hyperparameters()`.
It takes dots as its only argument, so we just pass the space of parameters
we want to search with hyperparameter names. To do so, we have three
functions available to us: `make_discrete()` takes discrete values
as arguments, `make_range()` takes a minimum and maximum value for a continuous
parameter, and `mixture()` takes multiple discrete or continuous
spaces and mixes them with user-specified weights. We then pass
the hyperparameter object into the `parameters` argument of a learner,
rather than passing them directly as we did before.

To demonstrate the capabilities of `enfold`, the last code block below
shows two types of grids for xgboost: one simple discrete grid without
a search engine (exhaustive search), and one grid with continuous parameters.
In the latter case, we optimize over penalty parameters `alpha` and `lambda`,
and use the `forbid()` method to ensure that only one of them can be larger
than zero. We also specify that they are sampled from a log-uniform distribution
mixed with a point mass at zero. We additionally set `n_snapshots` to 10
at an `nrounds` of 200, which makes xgboost become a list learner with
predictions output every 20 rounds without needing to re-train.

```{r xgboost grids, eval = TRUE}
# A simple search space
grid_simple <- specify_hyperparameters(
  max_depth = make_discrete(2, 4),
  alpha = make_discrete(0, 1)
)

# A complex search space
grid_complex <- specify_hyperparameters(
  max_depth = make_discrete(2, 3, 4),
  alpha = mixture(
    make_discrete(0),
    make_range(1e-5, 10, sample_space = sample_log_uniform()),
    weights = c(0.5, 0.5)
  ),
  lambda = mixture(
    make_discrete(0),
    make_range(1e-5, 10, sample_space = sample_log_uniform()),
    weights = c(0.5, 0.5)
  )
) |>
  forbid(alpha > 0 && lambda > 0)

# Training the simple grid and evaluating its performance
simple_xgb <- lrn_xgboost("XGB Simple", parameters = grid_simple)

simple_xgb_risk <- initialize_enfold(x = predictors, y = outcome) |>
  add_learners(simple_xgb) |>
  add_metalearners(mtl_selector("Select")) |>
  add_cv_folds(inner_cv = 3L, outer_cv = 3L) |>
  fit() |>
  risk_learners(loss_fun = loss_gaussian(), type = "cv")

# Training the complex grid and evaluating its performance
complex_xgb <- lrn_xgboost(
  "XGB Complex",
  parameters = grid_complex,
  nrounds = 200,
  n_snapshots = 10
)

# This wraps around our xgboost with the complex hyperparameter specification.
# make_score tells grd_early_stop() how to judge the performance of a list learner.
# Choosing selector here means only the best given a parameter configuration counts.
early_stop_complex_xgb <- grd_early_stop(
  complex_xgb,
  max_candidates = 30L,
  score = make_score(loss_gaussian(), mtl_selector("S"))
)

complex_xgb_risk <- initialize_enfold(x = predictors, y = outcome) |>
  add_learners(early_stop_complex_xgb) |>
  add_metalearners(mtl_selector("Select")) |>
  add_cv_folds(inner_cv = 3L, outer_cv = 3L) |>
  fit() |>
  risk_learners(loss_fun = loss_gaussian(), type = "cv")

# Now for performance evaluation:
# Simple grid:
simple_xgb_risk
setNames(
  min(simple_xgb_risk),
  names(which.min(simple_xgb_risk))
)

# The number of fit models is large, so we only show the risk of the best one
# Number of learners trained (maximum is max_candidates * n_snapshots):
length(complex_xgb_risk)

# Best one:
setNames(
  min(complex_xgb_risk),
  names(which.min(complex_xgb_risk))
)
```

## Session info

```{r session-info}
sessionInfo()
```
